Efficient FPGA Implementation of\\ Digit Parallel Online Arithmetic Operators

### 2014-07-01 FPT2014/bare_conf.tex
Georges comment 3_2

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components were originally designed to perform computations in digit serial with most significant digit first, resulting in the ability to chain arithmetic operators together for low latency. More recently, research has shown that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct RTL synthesis, the proposed architectures achieve slice savings of over $67\%$ and $69\%$, and speed-ups of over $1.2\times$ and $1.5\times$ for adders and multipliers, respectively. As a result, the area overheads of using online adders and multipliers in place of traditional arithmtic primitives is reduced from 
% Furthermore in comparison to standard arithmetic primitives, this translates to reduction of the area overheads from up to $8.41\times$ to $1.88\times$ for adders, and from $8.11\times$ to $1.84\times$ for multipliers using direct implementation, to $1.88\times$ and $0.54\times$ using our approaches.

@@

In summary, the main contributions of this paper are:
\begin{enumerate}
    \item Efficient methods to map online adders and multipliers to modern FPGAs;
    \item Discussion of how to implement correctly-rounded online multipliers for a chosen output precision;
    \item Demonstration of the performance improvements of online operators in terms of silicon area and operating frequency.
\end{enumerate}

@@

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
    \vspace{-6ex}
	\caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
    \vspace{2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$ \cite{Ercegovac_OnlineMult}.

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}
%
\begin{eqnarray}\label{Eq:SelFunc_OM}
\small
  sel(W_{[j]})=\begin{cases}
    1 & \text{ if } W_{[j]} \geqslant \frac{1}{2} \\
    0 & \text{ if } -\frac{1}{2}\leqslant W_{[j]}<\frac{1}{2} \\
    \overline{1} & \text{ if } W_{[j]}<-\frac{1}{2}
  \end{cases}
\normalsize
\end{eqnarray}

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on 4:2 compressors, as outlined in Fig.~\ref{Fig:PreviousWork} within the gray background. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the two FAs within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al \cite{FPT09RA} and Hormigo et al \cite{ASAP09RA} described very similar mapping techniques for online adders with SD representations and CS representations, respectively. They share the common idea of building the online adder based on the logic block (\texttt{LB}) as highlighted within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the two 3:2 compressors can be greatly reduced.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan2.pdf}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The gray background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\vspace{-1ex}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target FPGAs with 4-input LUTs (4-LUT) and two LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and four LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or wrong results. For instance, if two \texttt{LB}s are mapped to a slice with four LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} should not be used because the its carry input cannot be explicitly initialized. Due to the same reason, the output of \texttt{XORCY2} should not be used.

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork2.pdf}
	\vspace{-1ex}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    \vspace{-1ex}
\end{figure}

@@

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  We also compare these architectures against a ripple carry adder (\texttt{RCA}), which uses conventional arithmetic. While our experiments are targeting on Xilinx Virtex-6 FPGAs with speed grade -1, our methods are applicable to any FPGAs with the same slice architecture such as all the Xilinx 7 series FPGAs. Notice that in order to utilize the carry resources and avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.7. It can be seen that both designs with online arithmetic achieve relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay of the online adder is theoretically independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. More specifically, the frequency of \texttt{OA\_behv} is faster than \texttt{RCA} when $N\geqslant30$, whereas the proposed online adder outperforms \texttt{RCA} for $N>16$. As a result, this makes the \texttt{OA\_new} more attractive for a wider range of applications. Furthermore, the speed-ups of \texttt{OA\_new} against \texttt{OA\_behv} are over $1.2\times$.

We also compare the resource usage of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig~\ref{Fig:AdderArea}. We can see that in comparison to the original \texttt{OA\_behv}, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in slices. This is because the dedicated FPGA resources are fully utilized. These area savings reduce the overhead of an \texttt{OA} over an \texttt{RCA} from upto $2.95\times$ to $1.98\times$ for LUTs, and from $8.41\times$ to $1.88\times$ for slices. It is unlikely that the overhead can be completely avoided when using an \texttt{OA} because of the redundant number system. Nevertheless, with the reduced overhead, using \texttt{OA} to take advantage of the graceful degradation when operating beyond the deterministic clocking region may become more attractive for an FPGA developer.

@@

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $z_{j}  ~~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$					\vspace{.5ex}
  \end{algorithmic}
  \vspace{-1ex}
\end{algorithm}

@@

\subsection{FPGA Implementation}
We now present an area efficient implementation of Algorithm~\ref{Algorithm:OM_DigitParallel} on FPGAs. The general structure of a 4-digit \texttt{OM} using the proposed algorithm is shown in Fig.~\ref{Fig:PMStructure}(a), and the structure of a single stage $j$ is shown in Fig.~\ref{Fig:PMStructure}(b) with the word-lengths of all signals labelled. Area optimizations can be performed on all modules. It can be seen that in each stage, an $(N+2)$ digit online adder is used to derive $W_{[j]}$. Our proposed online adder architecture can be employed. The selection logic takes 3 input digits (6 bits) and generates 1 output digit (2 bits). Hence it can be implemented using two 6-LUTs for each output bit. Similarly the generation of $P_{[j]}$ can also be implemented using one 6-LUT, since only the integer digits of $W_{[j]}$ need to be modified due to the selection of $Z_{[j]}$. In addition, the structure of Stage 1 in Fig~\ref{Fig:PMStructure}(a) can be further optimized by removing the online adder, because $P_{[0]}=0$.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/ParallelMult_Structure.pdf}
	\caption{(a) Structure of a 4-digit online multiplier using the proposed algorithm. (b) Structure of one stage. The word-length of all signals are labelled in terms of the number of digits. $N$ denotes the word-length of the input signals.}
	\label{Fig:PMStructure}
	\vspace{-1ex}
\end{figure}

@@

The comparison of area in terms of the cost of LUTs and slices with respect to different operand word-lengths are shown in Fig.~\ref{Fig:OMArea}. The results are obtained from ISE 14.7 after mapping the design to the Virtex-6 FPGA. In comparison to \texttt{OM\_behv}, significant area savings have been obtained by the optimized design \texttt{OM\_full}. Specifically, the LUTs and slice utilization reduces $59.0\%\sim70.1\%$ and $68.6\%\sim81.7\%$, respectively. In comparison to the \texttt{CoreGen} multiplier, the area overhead of online multipliers drops significantly using the proposed new structure. In order to generate full precision product, the area overhead of \texttt{OM\_behv} against \texttt{CoreGen} are $3.94\times\sim5.82\times$ for LUTs, and $3.58\times\sim8.11\times$ for slices. Using the proposed structure \texttt{OM\_full} the area overheads drop to $1.48\times\sim1.83\times$ for LUTs and up to $1.84\times$ for slices. Furthermore as mentioned in the introduction, a traditional multiplier is unable to compute this correctly rounded output using less silicon area without extensive simulation. In contrast, using our approach we can trade area for precision and tune output precision to be the minimum desired. As an example, if only half of the output precision is required, \texttt{OM\_half} can save silicon area over the traditional \texttt{CoreGen} multiplier. The area saving varies from $40\%$ to $55\%$ for LUTs, and $46\%\sim56\%$ for slices.

@@

We comment that there remain a large gap in terms of frequency between the proposed online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This is because in the online multiplier, even with the carry resources in the FPGA being used efficiently, each 4-digit carry logic within a slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is hyperbole for large operand word-lengths. This also explains why \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, because the critical path of both designs are identical in theory.

>>>>

\begin{abstract}
%\boldmath
Online arithmetic has been widely studied for ASIC implementation. Online components were originally designed to perform computations in digit serial with most significant digit (MSD) first, resulting in the ability to chain arithmetic operators together for low latency. More recently, research has shown that digit parallel online operators can fail more gracefully when operating beyond the deterministic clocking region in comparison to operators with conventional arithmetic. Unfortunately, the utilization of online arithmetic operators in the past has required a large area overhead for FPGA implementation. In this paper, we propose novel approaches to implement the key primitives of online arithmetic, adders and multipliers, efficiently on modern Xilinx FPGAs with 6-input LUTs and carry resources. We demonstrate experimentally that in comparison to a direct RTL synthesis, the proposed architectures achieve slice savings of over $67\%$ and $69\%$, and speed-ups of over $1.2\times$ and $1.5\times$ for adders and multipliers, respectively. As a result, the area overheads of using online adders and multipliers in place of traditional arithmetic primitives is reduced from $8.41\times$ and $8.11\times$ to $1.88\times$ and $1.84\times$ respectively. Finally, because an online multiplier generates MSDs first, we also demonstrate the method to create an online multiplier with a reduced precision output that is smaller than a traditional multiplier producing the same result. We show that this can lead to silicon area savings of up to $56\%$.
% Furthermore in comparison to standard arithmetic primitives, this translates to reduction of the area overheads from up to $8.41\times$ to $1.88\times$ for adders, and from $8.11\times$ to $1.84\times$ for multipliers using direct implementation, to $1.88\times$ and $0.54\times$ using our approaches.

@@

In summary, the main contributions of this paper are:
\begin{enumerate}
    \item Efficient methods to map online adders and multipliers to modern FPGAs;
    \item A method to implement correctly-rounded online multipliers for a chosen output precision;
    \item Demonstration of the performance improvements of online operators in terms of area and frequency.
\end{enumerate}

@@

\subsection{Binary Online Addition}
Adders serve as a critical building block for arithmetic operations. To perform digit-parallel online addition, a redundant adder can be used directly. The structure of an online adder where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, with the cost of one extra FA for each digit of operands. This makes the online adder suitable for building up more complex arithmetic operators such as multipliers to accelerate the sum of partial products \cite{RedundantMult_1987}.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder2.pdf}
    \vspace{-6ex}
	\caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
    \vspace{-2ex}
	\label{Fig:Radix2SD_adder}
\end{figure}

@@

\subsection{Binary Online Multiplication}
Multiplication is another key arithmetic operator. Typically, online multiplication is performed in a recursive digit serial manner, as illustrated in Algorithm~\ref{Algorithm:OnlineMult} \cite{Ercegovac_OnlineMult} where both inputs and outputs are $N$-digit numbers as given in~(\ref{Eq:Online_Operands}). For a given iteration $j$, the product digit $z_j$ is generated MSD-first through a selection function $sel()$. For any radix $r$ and chosen digit set, there exits an appropriate selection method and a value of $\delta$ which ensure convergence. As the binary radix is used most commonly in computer arithmetic, we keep $r=2$ throughout this paper with the corresponding redundant digit set $\{\overline{1},0,1\}$. In this case $sel()$ is given by (\ref{Eq:SelFunc_OM}), and the selection is based on two integer digits and one fractional digit of $W_{[j]}$ \cite{Ercegovac_OnlineMult}.
%
\begin{eqnarray}\label{Eq:SelFunc_OM}
\small
  sel(W_{[j]})=\begin{cases}
    1 & \text{ if } W_{[j]} \geqslant \frac{1}{2} \\
    0 & \text{ if } -\frac{1}{2}\leqslant W_{[j]}<\frac{1}{2} \\
    \overline{1} & \text{ if } W_{[j]}<-\frac{1}{2}
  \end{cases}
\normalsize
\end{eqnarray}

\begin{algorithm}[tbp]
  \caption{Online Multiplication}  \label{Algorithm:OnlineMult}
  \begin{algorithmic}[1]
   \State \textbf{Initialization:}~$X_{[-\delta]}=Y_{[-\delta]}=P_{[-\delta]}=0$\vspace{.5ex}
    \For{$j=-\delta,~-\delta+1,~\cdots,~2N-1$}\vspace{.5ex}
        \State $H_{[j]} \leftarrow r^{-\delta}\left(x_{j+\delta+1}\cdot Y_{[j+1]}+y_{j+\delta+1}\cdot X_{[j]}\right)$\vspace{.5ex}
        \State $W_{[j]} \leftarrow P_{[j]} + H_{[j]}$\vspace{.5ex}
        \State $z_{j}~ \leftarrow sel(W_{[j]})$\vspace{.5ex}
        \State $P_{[j+1]} \leftarrow r\left(W_{[j]}-Z_{[j]}\right)$\vspace{.5ex}
    \EndFor
  \end{algorithmic}
\end{algorithm}
\vspace{-1ex}

@@

Specifically in type 1, both works took advantage of the built-in carry resources in FPGAs. Conventionally the ASIC implementation of online adders is based on 4:2 compressors, as outlined in Fig.~\ref{Fig:PreviousWork} within the gray background. However, directly applying this approach in the FPGAs could be less efficient. This is because there is no carry propagation between the two FAs within a 4:2 compressor, and the net delay between them can be large. Instead, Kamp et al \cite{FPT09RA} and Hormigo et al \cite{ASAP09RA} described very similar mapping techniques for online adders with SD representations and CS representations, respectively. They share the common idea of building the online adder based on the logic block (\texttt{LB}) as highlighted within the dotted circle in Fig.~\ref{Fig:PreviousWork}. In this case, the fast-carry logic in the FPGA can be employed, and the delay between the two 3:2 compressors can be greatly reduced.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/SDAdder_42comp_Spartan2.pdf}
	\vspace{-3ex}
	\caption{Map the online adder onto Spartan FPGAs using the fast-carry resources. The gray background highlights the 4:2 compressor. Dotted circle indicates the logic block (LB) which can be mapped to the FPGA using the carry resources.}
	\vspace{-4ex}
	\label{Fig:PreviousWork}
\end{figure}

However, we notice the major limitations of both approaches that they only target FPGAs with 4-input LUTs (4-LUT) and two LUTs within a logic slice, such as the Xilinx Spartan series and the Altera Cyclone series. This is naturally reasonable because one \texttt{LB} can be mapped to a single slice. Nevertheless, for FPGAs with 6-input LUTs (6-LUT) and four LUTs in a slice, such as the Xilinx Virtex series and all Xilinx 7 series FPGAs, directly applying these approaches will result in either resource waste or wrong results. For instance, if two \texttt{LB}s are mapped to a slice with four LUTs as seen in Fig.~\ref{Fig:PreWorkVirtex6}, the outputs of \texttt{LB1} should not be used because the its carry input cannot be explicitly initialized. Due to the same reason, the output of \texttt{XORCY2} should not be used.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.45\textwidth]{./Figures/SDadder_FastCarry_PreviousWork2.pdf}
	\vspace{-2ex}
	\caption{An example illustrating that direct applying approaches in \cite{FPT09RA} on a Virtex-6 FPGA will results in faulty outputs, because the carry input of \texttt{LB1} cannot be explicitly initialized.}
	\label{Fig:PreWorkVirtex6}
    \vspace{-2ex}
\end{figure}

@@

\subsection{Performance Analysis}\label{subsec:OA_Performance}
In our experiments, we compare the proposed online adder (\texttt{OA\_new}) against the original design which is implemented using functional descriptions in Verilog based on 4:2 compressors (\texttt{OA\_behv}).  We also compare these architectures against a ripple carry adder (\texttt{RCA}), which uses conventional arithmetic. While our experiments target on Xilinx Virtex-6 FPGAs (speed grade -1), our methods are applicable to any FPGAs with the same slice architecture such as all the Xilinx 7 series FPGAs. Notice that in order to utilize the carry resources and avoid logic optimizations by the synthesis tool, the Xilinx primitive component ``\texttt{Carry4}" is used to create \texttt{OA\_new}.

@@

We record the rated frequencies of all designs when varying the operand word-lengths. The results are shown in Fig.~\ref{Fig:AdderFreq}. The frequency values are obtained through Xilinx Timing Analyzer after placing and routing the designs in ISE 14.7. It can be seen that both designs with online arithmetic achieve relatively stable operating frequencies across a variety of operand word-lengths. This is as expected, because the critical path delay of the online adder is theoretically independent of the operand precision. In comparison, the rated frequency of the \texttt{RCA} drops dramatically for larger word-lengths. It follows that both online adders outperform \texttt{RCA} for large word-lengths. More specifically, the frequency of \texttt{OA\_behv} is faster than \texttt{RCA} when $N\geqslant30$, whereas our new online adder outperforms \texttt{RCA} for $N>16$. As a result, this makes the \texttt{OA\_new} more attractive for a wider range of applications. Furthermore, the speed-ups of \texttt{OA\_new} against \texttt{OA\_behv} are over $1.2\times$.

We also compare the resource usage of all three adder designs with respect to a variety of operand word-lengths. The results are presented in Fig.~\ref{Fig:AdderArea}. We can see that in comparison to the original \texttt{OA\_behv}, our implementation \texttt{OA\_new} achieves significant area savings: $25\%\sim 33\%$ in LUTs and $67\%\sim77\%$ in slices. This is because the dedicated FPGA resources are fully utilized. These area savings reduce the overhead of an \texttt{OA} over an \texttt{RCA} from $2.95\times$ to $1.98\times$ for LUTs, and from $8.41\times$ to $1.88\times$ for slices. It is unlikely that the overhead can be completely avoided when using an \texttt{OA} because of the redundant number system. Nevertheless, with the reduced overhead, using \texttt{OA} to take advantage of the graceful degradation when operating beyond the deterministic clocking region may become more attractive for an FPGA developer.

@@

\begin{algorithm}[tbp]
  \caption{Digit Parallel Online Multiplication}\label{Algorithm:OM_DigitParallel}
  \begin{algorithmic}[1]
    \State \textbf{Initialization:} $P_{[0]}=0$						\vspace{.5ex}
    \For{$j=1,~2,~\cdots,~N$}										\vspace{.5ex}
    	\State $Xy_j \leftarrow X \cdot y_j$						\vspace{.5ex}
    	\State $W_{[j]}    \leftarrow  P_{[j-1]} + Xy_j$			\vspace{.5ex}
     	\State $z_{j}  ~~      \leftarrow  sel(W_{[j]})$			\vspace{.5ex}
	 	\State $P_{[j+1]}  \leftarrow  r\left(W_{[j]}-Z_{[j]}\right)$	\vspace{.5ex}
    \EndFor															\vspace{.5ex}
    \State $Z_{[N+1:2N]} \leftarrow frac(W_{[N]})$					\vspace{.5ex}
  \end{algorithmic}
  \vspace{-1ex}
\end{algorithm}
\vspace{-1ex}

@@

\subsection{FPGA Implementation}
We now present an area efficient implementation of Algorithm~\ref{Algorithm:OM_DigitParallel} on FPGAs. The general structure of a 4-digit \texttt{OM} using the proposed algorithm is shown in Fig.~\ref{Fig:PMStructure}(a), and the structure of a single stage $j$ is shown in Fig.~\ref{Fig:PMStructure}(b) with the word-lengths of all signals labelled. Area optimizations can be performed on all modules. It can be seen that in each stage, an $(N+2)$ digit online adder is used to derive $W_{[j]}$. Our proposed online adder architecture can be employed. The selection logic takes 3 input digits (6 bits) and generates 1 output digit (2 bits). Hence it can be implemented using two 6-LUTs for each output bit. Similarly the generation of $P_{[j]}$ can also be implemented using one 6-LUT, since only the integer digits of $W_{[j]}$ need to be modified due to the selection of $Z_{[j]}$. In addition, the structure of Stage 1 in Fig~\ref{Fig:PMStructure}(a) can be further optimized by removing the online adder, because $P_{[0]}=0$.
%
\begin{figure}[tbp]
	\centering
	\includegraphics[width=.5\textwidth]{./Figures/ParallelMult_Structure.pdf}
	\vspace{-3ex}
	\caption{(a) Structure of a 4-digit online multiplier using the proposed algorithm. (b) Structure of one stage. The word-length of all signals are labelled in terms of the number of digits. $N$ denotes the word-length of the input signals.}
	\label{Fig:PMStructure}
	\vspace{-1ex}
\end{figure}

@@

The comparison of area in terms of the cost of LUTs and slices with respect to different operand word-lengths are shown in Fig.~\ref{Fig:OMArea}. The results are obtained from ISE 14.7 after mapping the design to the Virtex-6 FPGA. In comparison to \texttt{OM\_behv}, significant area savings have been obtained by the optimized design \texttt{OM\_full}. Specifically, the LUTs and slice utilization reduces $59.0\%\sim70.1\%$ and $68.6\%\sim81.7\%$, respectively. In comparison to the \texttt{CoreGen} multiplier, the area overhead of online multipliers drops significantly using the proposed new structure. In order to generate full precision product, the area overhead of \texttt{OM\_behv} against \texttt{CoreGen} are $3.94\times\sim5.82\times$ for LUTs, and $3.58\times\sim8.11\times$ for slices. Using the proposed structure \texttt{OM\_full} the area overheads drop to $1.48\times\sim1.83\times$ for LUTs and up to $1.84\times$ for slices. Furthermore as mentioned in the introduction, a traditional multiplier is unable to compute this correctly rounded output using less silicon area. In contrast, using our approach we can trade area for precision and tune output precision to be the minimum desired. As an example, if only half of the output precision is required, \texttt{OM\_half} can save silicon area over the traditional \texttt{CoreGen} multiplier. The area saving varies from $40\%$ to $55\%$ for LUTs, and $46\%\sim56\%$ for slices.

@@

We comment that there remain a large gap in terms of frequency between the proposed online multipliers and the \texttt{CoreGen} multiplier, especially for large operand word-lengths. This is because in the online multiplier, even with the carry resources in the FPGA being used efficiently, each 4-digit carry logic within a slice is divided into two parts, as shown previously in Section~\ref{Sec:OA_FPGA}. Therefore the critical path delay of an \texttt{OM} is determined by both the carry logic delays, and the net delay among different stages, which is huge for large operand word-lengths. This also explains why \texttt{OM\_full} and \texttt{OM\_half} run at similar frequencies, because the critical path of both designs are identical in theory.
